Truthfulness Verification System

### 2011-04-13 trunk/proposal/paper.tex
Rearranged

{\em Mar 29 to Apr 12}
\begin{itemize}
\item Extraction of meaningful entities(noun phrase) in topic unit and dount unit by checking all combination of content words
\item Ranking based on finding content words in the interesction of two relevant wiki articles. 
\item Extracting 
\end{itemize}

@@

\subsubsection{Use of Wikipedia}
\label{sec:UseOfWikipedia}
Although structured databases such as Freebase, DBPedia extract the information from Wikipedia and format them into a structured and query-friendly way, it does not contain all the relevant information all the time. For example, consider the following sentence:

\emph{Les Paul invented the electric guitar}

The wikipedia page "http://en.wikipedia.org/wiki/Electric\_guitar" contains the inventor of the electric guitar in one of the sentences in the body. But this information is not available in the Freebase.

We can use a very simple and elegant method to find the relevant information from Wikipedia by using simple web-scrapping and pattern matching using regular expression. The process is described below.

After getting the wikipedia page we can convert the html file into text file and then we can separate all the sentences in different lines. The sentence is the part of the whole text that exist between two full-stops. Another important observation is that the relevant information will be within the same sentence in the question-answer type sentences almost all the time. So we can match the content words to the sentences extracted from Wikipedia page and verify the truthfulness of the sentence.

For the example sentence above, after processing the Wikipedia page and matching by inventor we get following two sentences:
\begin {itemize}
\item \emph{"Commercial production began in late summer of 1932 by Electro-Patent-Instrument Company Los\_Angeles, a partnership of Adolph\_Rickenbacker, PaulBarth and George\_Beauchamp, the inventor"}

\item \emph{In a morecontemporary style, Little Willie Joe, the inventor of the Unitar, had a rhythmand\_blues instrumental hit in the 1950s with "Twitchy", recorded with the ReneHall Orchestra}
\end{itemize}

When we do a regular expression search of the $DU$ "Les Paul" in these two sentences, we do not get any match. Hence we can conclude that the statement is not true.

@@

\section{Ranking Algorithm 2}
The main idea behind this ranking algorithm for alternative units is that Wikipedia should contain content words from doubt statement in the same sentence because of the similarity in the context. Details description of the algorithm is given below.

@@

\subsubsection{Luna 2 is the first spacecraft on the moon}
The alternative units for this sentence are "Luna 2", "Apollo", "Land", "Landing", and "Nasa". Our algorithm returns true for both Luna 2 and Apollo and return false for all others. Now, both Luna 2 and Apollo are in reality true considering the facts that Luna 2 was the first unmanned spacecraft and Apollo was the first manned spacecraft. Hence, we can't say one of the Luna 2 or Apollo as false if farther information is not given. So algorithm 2 can produce correct result for this sentence also.

@@

\section{Conclusion}
It this paper, we described current system for truthfulness verification of statements found in the web called T-Verifier and described the possible improvements to the current system. We will primarily use authentic information sources such as Wikipedia and Ontologies extracted from Wikipedia such as FreeBase, DBPedia, and YOGA to found the truthfulness of doubtful statements.

>>>>

{\em Mar 29 to Apr 12}
See page~[\pageref{Mar29toApr12}] of section~\ref{sec:UseOfWikipedia}
\begin{itemize}
\item Extraction of meaningful entities(noun phrase) in topic unit and dount unit by checking all combination of content words
\item Ranking based on co-occurence of content words in the interesction of bag of words generated from wiki-articles. 
\item Ranking based on co-occurence of content words in the sentences generated from wiki-articles. 
\end{itemize}

@@

\section{Wikipedia API}

@@

\section{Use of Wikipedia}
\label{sec:UseOfWikipedia}
We have tried two approaches for using the Wikipedia. In both we utilize the content words to identify articles and there upon treat them separately. In the first approach we treat the articles as bag of words and do not consider the sentence structure and in the next approach we retain the logical structure of each sentence and try to infer from word overlaps.
%When we do a regular expression search of the $DU$ "Les Paul" in these two sentences, we do not get any match. Hence we can conclude that the statement is not true.

\label{Mar29toApr12}

\subsection{Bag of words}
\begin{enumerate}
\item Step I: Detection of wiki articles from $DU$, $AU$, $TU$.
\begin {enumerate}
\item For the $AU$s, we take the words as they appear in the output of T-verifier(mainly because most of them are single word). For $TU$ we use a basic exponential algorithm that takes all possible combinations of the content words.
\item The default configuration of T-verifier generates a total of 5 $AU$s, without any ranking.  For each content word in a $AU$, or a combination of words in case the $AU$ consists of multiple content words, we check if there exists a wikipedia article. There may be three cases:
\begin{enumerate}
\item No article. Continue with next word or word combination. 
\item One article. Go to next step
\item More than one article. The page lands into a "disambiguation" page. We discuss this later. 
\end{enumerate}
\end{enumerate}
\item We remove the html tags from the pages, and convert them into simple bag of words. Repetation of a word if any is retained.
\item We select the following pairwise combination - a wikiarticle detected from $AU$, ($W_{au_{i}}$) and a wikiarticle detected from $TU$ ($W_{tu_{j}}$).
\item We take the common words $c_{ij} = {w_{1}, w_{2}, \ldots}$ matched by regular expressions in these two bags.
\item Since this is a regular expression matching we will have two different values for $c_{ij}$ = common($W_{au_{i}}$, $W_{tu_{j}}$) and $c_{ji}$ = common($W_{tu_{j}}$, $W_{au_{i}}$). This is explained below.
\item In this collection of common words, $c_{ij}$, we search each of the content words of the $TU$, not included in the title of $W_{tu_{j}}$ 
\item We observe the maximum number of matches in all of such interesctions and use the count to rank the $AU$-s generated by T-verifer. 
\end{enumerate}

\paragraph{Handling Disambiguation}
For the testing that we have done with the five problematic sentences, we manually first used the disambiguation page itself in the above process. But as expected, it did not yeild good results. We are experimenting with the following ideas for disabmigutation.
\begin{enumerate}
\item Apply the above steps for each of the page linked from the disambiguation page and get the one which gives the maximum count
\item Use the heading under which it appears in the disambiguation page, the one sentence summary appearing adjacent to the entry in the disambiguation page, and the "Categories" under which a particular article has been tagged.
\end{enumerate}
For this report, we manually checked the disambiguation pages by searching them for content words. This significantly improved the results.

\paragraph{Word matching}
We found that the intersection of the two bags of words, $W_{au_{i}}$ and $W_{tu_{i}}$ produced different results based on the ordereing of the articles while performing the intersection operation. This is because the regular expression match finds the maximal match. Lets consider that the first file, $W_{au_{i}}$ has "invent", and the second file, $W_{tu_{i}}$ has "invented", and "inventor". So when we pick a word in $W_{au_{i}}$ and try to match words in $W_{tu_{i}}$, we will find 2 matches. However, the words from second file, both "invented", and "inventor" will not find any matches in ($W_{au_{i}}$). Taking both way intersections will ensure we do not miss out different morphological variations of the same word appearing in the context common to the both articles.

\paragraph{Example}

A description of the above steps applied to the five problematic sentences
\begin{enumerate}
\item Sentence 1: Les Paul invented the electric guitar.
\begin{enumerate}
\item AU: Les\_Paul, Gibson Les, Leo, Llyod, Adolph Rickenbacker. 
\item Wiki articles detected from AU($W_{au_{i}}$): Les\_Paul, Leo (disambiguation page), Adolph\_Rickenbacker 
\item Wiki articles detected from TU($W_{tu_{i}}$): The\_guitar, The\_Electric\_guitar, Electric\_guitar, invented Invention 
\end{enumerate}

\item Sentence 2: Tom Hanks is the lead actress of the movie Sleepless in Sattle.
\begin{enumerate}
\item AU: hanks meg, meg ryan, tom hanks, nora, sam
\item Wiki articles detected from AU($W_{au_{i}}$): Meg\_Ryan, Tom\_Hanks, Nora(Disambiguation), Sam(Disambiguation page) 
\item Wiki articles detected from TU($W_{tu_{i}}$): Sleepless\_in\_Seattle, Lead, Lead-in, Leading\_actor(automatically redirected from Leading actress), Movie(automatically redirected from Film), Sleepless(disambiguation page), Seattle(lands on the correct page, though there is a disambiguation page).
\end{enumerate}

\item Sentence 3: Apollo was the first spacecraft on the moon.
\begin{enumerate}
\item AU: apollo, land, landing, luna
\item Wiki articles detected from AU($W_{au_{i}}$): Apollo, Land(disambiguation page), Landing(disambiguation page), Luna(disambiguation)
\item Wiki articles detected from TU($W_{tu_{i}}$): Fisrt(disambiguation page), Moon, Spacecraft.
\end{enumerate}

\item Sentence 4: english is the primary language of the philippines.
\begin{enumerate}
\item AU: english, filipino, spansih, education, history
\item Wiki articles detected from AU($W_{au_{i}}$): English disambiguation page), Filipino(disambiguation page), Spanish(disambiguation), Education, History
\item Wiki articles detected from TU($W_{tu_{i}}$): Primary (disambiguation page), language(disambiguation page), First\_Language (automatically redirected from Primary Languge), Philippines.

\end{enumerate}

\item Sentence 5: Michael Phelps is the fastest swimmer in the world
\begin{enumerate}
\item AU: michael, long, ernest, sullivan, alexander
\item Wiki articles detected from AU($W_{au_{i}}$): Michael, Alexander, Sullivan(disambiguation page), Long(disambiguation page), Ernest(disambiguation page) 
\item Wiki articles detected from TU($W_{tu_{i}}$): Fastest, Swimming(sport) (automatically redirected from Swimmer),World.

\end{enumerate}

\end{enumerate}

Please refer to the hand out analysis of the results.

\subsection{Sentence Splitting}

The main idea behind this ranking algorithm for alternative units is that Wikipedia should contain content words from doubt statement in the same sentence because of the similarity in the context. Details description of the algorithm is given below.

@@

%On reading the articles related above, there is no information that Rickenbacker actually inventing the electric guitar. He was a founding member of the Rickenbacker company that produces electric guitar. Les Paul on the other hand was a pioneer in designing and developing of what a prototype. Moreover, some of the categories attached with the Les Paul page are American musical instrument makers, Guitar makers, Inventors of musical instruments, National Inventors Hall of Fame inductees. In comparison Adolph Rickenbaker page has only one guitar related category - Guitar stubs, meaning the article is stuck in content.
\subsubsection{Luna 2 is the first spacecraft on the moon}
The alternative units for this sentence are "Luna 2", "Apollo", "Land", "Landing", and "Nasa". Our algorithm returns true for both Luna 2 and Apollo and return false for all others. Now, both Luna 2 and Apollo are in reality true considering the facts that Luna 2 was the first unmanned spacecraft and Apollo was the first manned spacecraft. Hence, we can't say one of the Luna 2 or Apollo as false if farther information is not given. So algorithm 2 can produce correct result for this sentence also.

@@

%After getting the wikipedia page we can convert the html file into text file and then we can separate all the sentences in different lines. The sentence is the part of the whole text that exist between two full-stops. Another important observation is that the relevant information will be within the same sentence in the question-answer type sentences almost all the time. So we can match the content words to the sentences extracted from Wikipedia page and verify the truthfulness of the sentence.

%For the example sentence above, after processing the Wikipedia page and matching by inventor we get following two sentences:
%\begin {itemize}
%\item \emph{"Commercial production began in late summer of 1932 by Electro-Patent-Instrument Company Los\_Angeles, a partnership of Adolph\_Rickenbacker, PaulBarth and George\_Beauchamp, the inventor"}

%\item \emph{In a morecontemporary style, Little Willie Joe, the inventor of the Unitar, had a rhythmand\_blues instrumental hit in the 1950s with "Twitchy", recorded with the ReneHall Orchestra}
%\end{itemize}

%\section{Conclusion}
%It this paper, we described current system for truthfulness verification of statements found in the web called T-Verifier and described the possible improvements to the current system. We will primarily use authentic information sources such as Wikipedia and Ontologies extracted from Wikipedia such as FreeBase, DBPedia, and YOGA to found the truthfulness of doubtful statements.
